### 2024-04-16 Cogsci_version_raw.Rmd
revised submission Cogsci, main article not yet including polished analytical argument

(ref:simulations) **A** Shows the results of simulations in a categorical setting with 3 choice options. Points represent average accuracy (left)/competence (right) values by degree of convergences (measured by the share of votes for an options), for different population competence distributions and sample sizes. **B** Shows the results in a continuous setting. Regression lines represent the correlation between accuracy (left)/competence (right) and convergence (measured by squared distance to true mean and reversed such that greater convergence corresponds to being more right on the x-axis). **C** Shows the different population competence distributions we considered in our simulations. In the continuous simulations, competence values of 0 corresponds to a very large standard deviation (1000, with a mean of 1500, on a scale from 1000 to 2000), thereby practically taking the form of a uniform distribution, while competence of 1 corresponds to a standard deviation of 1. In the categorical simulations, a competence value of 0 corresponds to chance (e.g. in a 3-choice-options scenario, an individual picking the correct answer with a probability of $1/3$), while competence of 1 corresponds to definitely picking the correct answer.

@@

# common plot
patchwork <-  plots_categorical / plots_continuous / (plot_competence_distributions(simulation_numeric) + plot_spacer())

>>>>

(ref:simulations) **A** Shows the different population competence distributions we considered in our simulations. In the continuous simulations, competence values of 0 corresponds to a very large standard deviation (1000, with a mean of 1500, on a scale from 1000 to 2000), thereby practically taking the form of a uniform distribution, while competence of 1 corresponds to a standard deviation of 1. In the categorical simulations, a competence value of 0 corresponds to chance (e.g. in a 3-choice-options scenario, an individual picking the correct answer with a probability of $1/3$), while competence of 1 corresponds to definitely picking the correct answer. **B** Shows the results of simulations in a categorical setting with three choice options. Points represent average accuracy (left)/competence (right) values by degree of convergences (measured by the share of votes for an options), for different population competence distributions and sample sizes. **C** Shows the results in a continuous setting. Regression lines represent the correlation between accuracy (left; measured by squared distance to true mean and reversed such that greater accuracy corresponds to top of the figure) or competence (right), respectively, and convergence (reversed such that greater convergence corresponds to being more right on the x-axis).

@@

# common plot
patchwork <- (plot_competence_distributions(simulation_numeric) + plot_spacer()) / 
  plots_categorical /
  plots_continuous

### 2024-04-16 README.qmd
revised submission Cogsci, main article not yet including polished analytical argument

- `appendix*` files contain the appendices that are automatically added `paper.Rmd` when rendering the file.
- `Cogsci*` files are documents used to generate a shorter version of our paper submitted for the CogSci 2024 conference. 
- `Experiment*` folders contain all files related to the respective experiment. The `/data` folders contain the raw data (`qualtrics.csv`) and cleaned data (`cleaned.csv`). The cleaning can be reproduced by running the `cleaning.Rmd`. `Experiment_1` and `Experiment_4` also contain the code for the simulations (`model.Rmd`) and several data files with simulated data. 
- `functions/` folder contains functions that we rely upon in our analysis scripts. To keep the `paper.Rmd` document neat and readable, we wrote all lengthy code bits as external functions.
- figures includes images used in the `paper.Rmd` file
- `howwise_0131.pdf` is the final submission to the CogSci 2024 conference, based on the raw files but edited in [overleaf](https://www.overleaf.com/). 
- `output` folder lists all figures generated by the `paper.Rmd` file as `.png` and `.pdf` files.
- `paper.pdf` is the rendered version of `paper.Rmd`
- `references.bib` contains the references
- `stimuli` folder contains all stimuli

>>>>

- `appendix*` files contain the appendices that are automatically added `paper.Rmd` when rendering the file.
- `Cogsci*` files are documents used to generate the initial submission of a shorter version of our paper for the CogSci 2024 conference.
- `Experiment*` folders contain all files related to the respective experiment. The `/data` folders contain the raw data (`qualtrics.csv`) and cleaned data (`cleaned.csv`). The cleaning can be reproduced by running the `cleaning.Rmd`. `Experiment_1` and `Experiment_4` also contain the code for the simulations (`model.Rmd`) and several data files with simulated data. 
- `functions/` folder contains functions that we rely upon in our analysis scripts. To keep the `paper.Rmd` document neat and readable, we wrote all lengthy code bits as external functions.
- figures includes images used in the `paper.Rmd` file
- `howwise_0131_initial_submission.pdf` is the initial, and `howwise_0131.pdf` the final submission (after review) to the CogSci 2024 conference, based on the raw files but edited in [overleaf](https://www.overleaf.com/). 
- `output` folder lists all figures generated by the `paper.Rmd` file as `.png` and `.pdf` files.
- `paper.pdf` is the rendered version of `paper.Rmd`
- `references.bib` contains the references
- `stimuli` folder contains all stimuli

### 2024-04-16 paper.Rmd
revised submission Cogsci, main article not yet including polished analytical argument

abstract: |
  Are people who agree on something more likely to be right and competent? Evidence suggests that people tend to make this inference. However, standard wisdom of crowds approaches only provide limited normative grounds . Using simulations and analytical arguments, we argue that when individuals make independent and unbiased estimates, under a wide range of parameters, individuals whose answers converge with each other tend to have more accurate answers and to be more competent. In 6 experiments (UK participants, total N = 1197), we show that participants infer that informants who agree have more accurate answers and are more competent, even when they have no priors, and that these inferences are weakened when the informants were systematically biased. In conclusion, we speculate that  inferences from convergence to accuracy and competence might help explain why people deem scientists competent, even if they have little understanding of science.

@@

There are two ways for answers to converge: when the answers are continuous, they can have a smaller range, or a lower variance; when the answers are categorical (i.e. there are only a few options), more answers can be the same. These two cases have been studied with different paradigms.

@@

There is clear evidence that, both in the continuous and the categorical case, participants infer that convergent opinions are more likely to be correct. There is tentative evidence that participants also infer that individuals who hold more convergent opinions are more likely to be competent. Would that second inference be justified? More precisely, if we consider individuals who have convergent opinions, by contrast with individuals whose opinions are less convergent, or whose opinion diverges from that of a majority, are we justified in thinking that the individuals with more convergent opinions are more likely to be right not only for this specific question, but that they would also be right for other (related) questions? We first provide a general analytic answer, before presenting simulations for the continuous, and then the categorical case.

@@

In the continuous case, we imagine a population of informants. The answers of each informant are drawn from a normally distributed random variable, centered on the correct answer (i.e., informants have no systematic bias). The variance of this distribution represents the individuals’ competence: the larger the variance, the lower the competence. If we observe a sample of this population, the sample mean (i.e. the mean of the answers of all the informants) is then the best estimate of the correct answer, and the sample variance (i.e. mean squared distance between the answers and the sample mean) is the best estimate of the population’s average competence (best is understood here as the least volatile estimator). This means that a population of informants whose answers converge more (lower sample variance) is, on average, more competent (informants tend to answer with a lower variance).

We can extend this argument from populations to individuals: Consider that competence varies within a population, such that each informant's answer is drawn from their own individual distribution--always centered on the correct answer, but with different variances. In this case, the distance between each informant's answer and the sample mean provides the best estimate of that informant's competence (i.e. individuals whose answers are further away from the sample mean tend to be less competent).

In the categorical case, we assume that informants answer a question with several categorical answers. We define the competence of an informant as the probability of choosing the correct answer. Due to the law of large numbers, the relative size of the majority--the share of informants who choose the most chosen answer--is, when the population is large enough, a good approximation of the average competence of the population (e.g. if the average competence is .66, and there are enough informants, then approximately 66% of informants will select the right answer). This relationship still exists, but is more noisy for small populations. This means that, at the population level, the more the answers converge, the more the population is, on average, competent (and the larger the population, the stronger the inference).

To make inferences on individual informants' competence on the basis of whether or not their answers converge with those of the majority in the sample, we must assume some distribution of competence in the population (assuming, as above, that the true average population competence is equal to the observed majority size in the sample). Using Bayes theorem, we can then show (see Electronic Supplementary Materials, ESM, section \@ref(analytical)) that the more informants agree with a focal informant, the more the focal informant can be inferred to be competent (Fig. \@ref(fig:figure-Benoit-2)). Additionally, the more competence varies in the population, the more the degree of convergence is informative of an informant's competence. It makes sense: if competence is roughly uniform in the population, then being part of a minority likely reflects bad luck, rather than incompetence. More generally, the fact that the focal individual is right (their 'accuracy') can be inferred more strongly than their competence, as there is noise in the answer choice (an incompetent individual can always pick the correct answer by chance and vice versa). Figure \@ref(fig:figure-Benoit-2) provides an example of a specific categorical choice, for two different distributions of competence.

@@

In order to better understand the influence of different parameters (e.g. degree of convergence, number of individuals) on the relationship between convergence, accuracy, and competence of informants, we conducted simulations.

@@

Groups of agents, with each agent’s competence varying, provide numerical answers. We measure how accurate these answers are, and how much they converge (i.e. how low their variance is). We then look at the relationship between convergence and both the accuracy of the answers and the competence of the agents.

@@

(ref:simulations) **A** Shows the different population competence distributions we considered in our simulations. In the continuous simulations, competence values of 0 corresponds to a very large standard deviation (1000, with a mean of 1500, on a scale from 1000 to 2000), thereby practically taking the form of a uniform distribution, while competence of 1 corresponds to a standard deviation of 1. In the categorical simulations, a competence value of 0 corresponds to chance (e.g. in a 3-choice-options scenario, an individual picking the correct answer with a probability of $1/3$), while competence of 1 corresponds to definitely picking the correct answer. **B** Shows the results of simulations in a categorical setting with three choice options. Points represent average accuracy (left)/competence (right) values by degree of convergences (measured by the share of votes for an options), for different population competence distributions and sample sizes. **C** Shows the results in a continuous setting. Regression lines represent the correlation between accuracy (left; measured by squared distance to true mean and reversed such that greater accuracy corresponds to top of the scale) or competence (right), respectively, and convergence (reversed such that greater convergence corresponds to being more right on the x-axis).

@@

All experiments were preregistered. Preregistration documents, data and code can be found on Open Science Framework [project page](https://osf.io/6abqy/) (<https://osf.io/6abqy/>). All analyses were conducted in R (version 4.2.2) using R Studio. For most statistical models, we relied on the `lme4` package and its `lmer()` function.

@@

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: "We will show you three financial advisors who are giving recommendations on investment decisions. They can choose between three investment options. Their task is to recommend one. You will see several such situations. They are completely unrelated: it is different advisors evaluating different investments every time. At first you have no idea how competent the advisors are: they might be completely at chance, or be very good at the task. It's also possible that some are really good while others are really bad. Some tasks might be difficult while others are easy. Your task will be to evaluate the performance of one of the advisors based on what everyone's answers are." To assess perceptions of accuracy, we asked: "What do you think is the probability of advisor 1 making the best investment recommendation?". Participants answered with a slider on a scale from 0 to 100. To assess perceptions of accuracy, we asked: "How competent do you think advisor 1 is regarding such investment recommendations?" Participants answered on a 7-point Likert scale (from (1)"not competent at all" to (2)"extremely competent").

@@

On the whole, participants thus appear to draw normatively justified inferences: (i) they infer that more convergent answers are more accurate; (ii) they infer that more convergent answers are more competent informants; (iii) inference (ii) tends to be stronger than inference (i); (iv) both inference (i) and (ii) are weaker when the informants have a conflict of interest, but not when they are merely discussing with each other. The only exception appears to be that participants do not take into account the number of options in categorical choices scenarios.

@@

The results--both simulations and experiments--are a novel contribution to the wisdom of crowds literature. In this literature--in particular that relying on the Condorcet Jury Theorem--a degree of competence is assumed in the individuals providing some answers. From that competence, it can be inferred that the individuals will tend to agree, and that their answers will tend to be accurate. Here we show that the reverse inference--from agreement to competence--is also warranted, and that it is warranted under a wide range of circumstances. We also show that participants, by and large, are able to draw rational inferences, inferring accuracy and competence from an observation of convergence, and doing so more when there are no alternative explanations for the observed convergence.

People might draw this inference in a variety of contexts, but the most prominent one might be science. Science is, arguably, the institution in which individuals end up converging the most in their opinions. For instance, scientists within many disciplines agree on things ranging from the distance between the solar system and the center of the galaxy to the atomic structure of DNA. This represents an incredible degree of convergence. When people hear that scientists have measured the distance between the solar system and the center of the galaxy, if they assume that there is a broad agreement within the relevant experts, this should lead them to infer that this measure is accurate, and that the scientists who made it are competent. Experiments have already shown that increasing the degree of perceived consensus among scientists tends to increase acceptance of the consensual belief [@vanstekelenburgScientificConsensusCommunicationContested2022], but it hasn't been shown yet that the degree of consensus also affects the perceived competence of scientists.

@@

The current study has a number of limitations. If the very abstract materials allow us to remove most of the priors that participants might have, they might also reduce the ecological validity of the experiments. Although the main results replicate well, and we can thus be reasonably certain of their robustness with the present samples, it's not clear how much they can be generalized. Experimental results with convenience samples can usually be generalized at least to the broader population the samples were drawn from (here, UK citizens) [@coppockGeneralizingSurveyExperiments2019]. However, we do not know whether they would generalize to other cultures.

>>>>

abstract: |
  Are people who agree on something more likely to be right and competent? Evidence suggests that people tend to make this inference. However, standard wisdom of crowds approaches only provide limited normative grounds. Using simulations and analytical arguments, we argue that when individuals make independent and unbiased estimates, under a wide range of parameters, individuals whose answers converge with each other tend to have more accurate answers and to be more competent. In 6 experiments (UK participants, total N = 1197), we show that participants infer that informants who agree have more accurate answers and are more competent, even when they have no priors, and that these inferences are weakened when the informants were systematically biased. In conclusion, we speculate that  inferences from convergence to accuracy and competence might help explain why people deem scientists competent, even if they have little understanding of science.

@@

The literature on the wisdom of crowds has treated separately situations with continuous answers, such as the weight of an ox in Galton’s famous observation [@galtonVoxPopuli1907], and with categorical answers, as when voters have to choose between two options, in the standard Condorcet Jury Theorem [@decondorcetEssaiApplicationAnalyse2014]. The continuous and the categorical case are typically modeled with different tools, and they have usually been studied in different empirical literatures (see below). Given that they both represent common ways for answers to converge more or less (e.g. when people give numerical estimates vs. vote on one of a limited number of options), we treat them both here, with different simulations and experiments.

@@

There is clear evidence that, both in the continuous and the categorical case, participants infer that convergent opinions are more likely to be correct. There is also tentative evidence that participants infer that individuals holding more convergent opinions are more likely to be competent. Is that second inference justified? More precisely, is it normatively warranted to claim that individuals holding convergent opinions are more likely to be right not only for this specific question, but also be right for other (related) questions? We first provide a general analytic answer, before presenting simulations for the continuous and categorical cases.

@@

In the continuous case, let us imagine a population of informants. Individual opinions are drawn from a normal distribution, centered on the correct answer [i.e., informants have no systematic bias, as for instance in Galton’s classic demonstration, @galtonVoxPopuli1907]. The variance of this distribution represents the individuals’ competence: the larger the variance, the lower the competence.  We observe the individual answers. In this setting, it is well known in statistics (see also Electronic Supplementary Materials, ESM) that the sample mean (i.e. the mean of the answers of all the informants) is the best estimator of the correct answer, and the sample variance (i.e. the mean squared distance between the answers and the sample mean) is the best estimator of the population’s average competence (best understood here as the least volatile estimator). This means that a population of informants whose answers converge more (lower sample variance) is, on average, more competent (informants tend to answer with a lower variance).

We can extend this argument from populations to individuals: Consider that competence varies within a population, such that each informant's answer is drawn from their own distribution -- always centered on the correct answer, but with different variances. In this case, the distance between each informant's answer and the sample mean provides the best estimate of that informant's competence (i.e. individuals whose answers are further away from the sample mean tend to be less competent).

In the categorical case, we define the competence of an informant as the probability of choosing the correct answer. The law of large numbers implies that the relative size of the majority – the share of informants who choose the most chosen answer – is, when the population is large enough, a good approximation of the average competence of the population (e.g. if the average competence is .66, and there are enough informants, then approximately 66% of informants will select the right answer). For smaller populations, the relationship holds, with some degree of noise. In other words, the more the answers converge, the more the population can be inferred to be competent (and the larger the population, the more reliable the inference).

Now, can we infer that informants belonging to the majority are more competent? To do so, we must assume some distribution of competence in the population. Using Bayes theorem, we can then show (see ESM, section \@ref(analytical)) that the more informants agree with a focal informant, the more the focal informant can be inferred to be competent (Fig. \@ref(fig:figure-Benoit-2)). We also find that , the more competence varies in the population, the more the degree of convergence is indicative of an informant’s competence. This reflects the fact that, if competence is roughly uniform in the population, then being part of a minority likely reflects bad luck, rather than incompetence. More generally, the fact that the focal individual is right (their ‘accuracy’) can be inferred more strongly than their competence, as there is noise in the answer choice (an incompetent individual can always pick the correct answer by chance and vice versa) (see ESM). Figure \@ref(fig:figure-Benoit-2) provides an example with a specific categorical choice scenario, under two different population distributions of competence.

@@

In order to better understand the influence of different parameters (e.g. degree of convergence, number of individuals) on the relationship between convergence, accuracy, and competence of informants, we conducted simulations. In all simulations, we assume agents to be unbiased and independent in their answers, but varying in their competence.

@@

Groups of agents, with each agents' competence varying, provide numerical answers. We measure how accurate these answers are, and how much they converge (i.e. how low their variance is). We then look at the relationship between convergence and both the accuracy of the answers and the competence of the agents.

@@

(ref:simulations) **A** Shows the different population competence distributions we considered in our simulations. In the continuous simulations, competence values of 0 correspond to a very large standard deviation (1000, with a mean of 1500, on a scale from 1000 to 2000), thereby practically taking the form of a uniform distribution, while competence of 1 corresponds to a standard deviation of 1. In the categorical simulations, a competence value of 0 corresponds to chance (e.g. in a 3-choice-options scenario, an individual picking the correct answer with a probability of $1/3$), while competence of 1 corresponds to definitely picking the correct answer. **B** Shows the results of simulations in a categorical setting with three choice options. Points represent average accuracy (left)/competence (right) values by degree of convergences (measured by the share of votes for an options), for different population competence distributions and sample sizes. **C** Shows the results in a continuous setting. Regression lines represent the correlation between accuracy (left; measured by squared distance to true mean and reversed such that greater accuracy corresponds to top of the figure) or competence (right), respectively, and convergence (reversed such that greater convergence corresponds to being more right on the x-axis).

@@

All experiments were preregistered. Preregistration documents, data and code can be found on Open Science Framework [project page](https://osf.io/6abqy/) (<https://osf.io/6abqy/>). All analyses were conducted in R (version 4.2.2) using R Studio. For most statistical models, we relied on the `lme4` package and its `lmer()` function. Unless mentioned otherwise, we report unstandardized model coefficients that can be interpreted in units of the scales we use for our dependent variables.

@@

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: "We will show you three financial advisors who are giving recommendations on investment decisions. They can choose between three investment options. Their task is to recommend one. You will see several such situations. They are completely unrelated: it is different advisors evaluating different investments every time. At first you have no idea how competent the advisors are: they might be completely at chance, or be very good at the task. It's also possible that some are really good while others are really bad. Some tasks might be difficult while others are easy. Your task will be to evaluate the performance of one of the advisors based on what everyone's answers are." To assess perceptions of accuracy, we asked: "What do you think is the probability of advisor 1 making the best investment recommendation?". Participants answered with a slider on a scale from 0 to 100. To assess perceptions of competence, we asked: "How competent do you think advisor 1 is regarding such investment recommendations?" Participants answered on a 7-point Likert scale (from (1)"not competent at all" to (2)"extremely competent").

@@

On the whole, participants thus appear to draw normatively justified inferences: (i) they infer that more convergent answers are more accurate; (ii) they infer that more convergent answers are coming from more competent informants; (iii) inference (i) tends to be stronger than inference (ii); (iv) both inference (i) and (ii) are weaker when the informants have a conflict of interest, but not when they are merely discussing with each other. The only exception appears to be that participants do not take into account the number of options in categorical choices scenarios.

@@

The results--both simulations and experiments--are a novel contribution to the wisdom of crowds literature. In this literature--in particular that relying on the Condorcet Jury Theorem--a degree of competence is assumed in the individuals providing some answers. From that competence, it can be inferred that the individuals will tend to agree, and that their answers will tend to be accurate. Here, we show that the reverse inference--from agreement to competence--is also warranted, and that it is warranted under a wide range of circumstances: If one does not suspect any systematic bias, convergence alone can be sufficient in determining who tends to be an expert. This finding qualifies work suggesting that people need a certain degree of expertise themselves, in order to figure out who is an expert [@nguyenCognitiveIslandsRunaway2020; @hahnHowGoodYour2018].

People might draw these inferences from convergence in a variety of contexts, but the most prominent one might be science. Science is, arguably, the institution in which individuals end up converging the most in their opinions. For instance, scientists within many disciplines agree on things ranging from the distance between the solar system and the center of the galaxy to the atomic structure of DNA. This represents an incredible degree of convergence. When people hear that scientists have measured the distance between the solar system and the center of the galaxy, if they assume that there is a broad agreement within the relevant experts, this should lead them to infer that this measure is accurate, and that the scientists who made it are competent. Experiments have already shown that increasing the degree of perceived consensus among scientists tends to increase acceptance of the consensual belief [@vanstekelenburgScientificConsensusCommunicationContested2022], but it hasn't been shown yet that the degree of consensus also affects the perceived competence of scientists.

@@

The current study has a number of limitations. In our simulations, we assume agents to be independent. Following previous work generalizing the Condorcet Jury Theorem to cases of informational dependency [@ladhaCondorcetJuryTheorem1992], more robust simulations would show that–while still assuming no systematic bias–our results hold even when agents influence each others' answers. If the very abstract materials allow us to remove most of the priors that participants might have, they might also reduce the ecological validity of the experiments. Although the main results replicate well, and we can thus be reasonably certain of their robustness with the present samples, it's not clear how much they can be generalized. Experimental results with convenience samples can usually be generalized at least to the broader population the samples were drawn from (here, UK citizens) [@coppockGeneralizingSurveyExperiments2019]. However, we do not know whether they would generalize to other cultures.
